Abstraction and Invariance for Algebraically Indexed Types

### 2012-07-09 paper/section-monoid.tex
Fixes to the monoid section to account for the restrictions on non-definability results

For the sets of relational environments, we use %make use of 
the
construction in \autoref{sec:constr-rel-env}, setting the chosen model
$\mathcal{M}$ to be the centre $Z(M)$ of $M$. % (recall that the
%centre of a monoid is the set of all elements that commute with every
%other element). 
Setting $R_{\tyPrimNm{M}}(m) = \{ (n,mn) \sepbar n
\in M \}$ and $R^\bullet_{\tyPrimNm{M}} = \emptyset$
%By setting the model $\mathcal{M}$ to be only the elements of $M$ that
%commute with everything else, we ensure that the following lemma
%holds:
gives:
\begin{lemma}
  For all $\Delta$ and $\rho \in \relEnv{E}_\Mon(\Delta)$,
  $(\eta_\Mon,\eta_\Mon) \in \rsem{\Gamma_\Mon}{\relEnv{E}}\rho$.
\end{lemma}

@@

\begin{theorem}
  \label{thm:monoid-indefinability}
  Let $m$ and $n$ be natural numbers. For all $n$-by-$m$-matricies of
  natural numbers $A$ and $m$-by-$1$ matricies of natural numbers $b$,
  there exists a program:
  \begin{equation}
    ; \Gamma_\Mon \vdash M : \forall
    i_1,\dots,i_m\mathord:\mathsf{M}.\ \tyPrim{M}{e_1} \to \dots \to
    \tyPrim{M}{e_n} \to \tyPrim{M}{e}
  \end{equation}
  where $e_j = i_1^{A_{j1}}\dots i_m^{A_{jm}}$ and $e = i_1^{b_1}\dots
  i_m^{b_m}$ iff there is a solution $\vec{x} = (x_1,...,x_n)$ in the
  natural numbers to $A x^\top = b$.
\end{theorem}

\begin{example}
  By \thmref{thm:monoid-indefinability}, it is impossible to write a
  program with the type $\forall m_1,m_2 \mathord: \mathsf{M}.\
  \tyPrim{M}{m_1m_2} \to \tyPrim{M}{m_1}$, and is thus % shows that it is
  impossible to convert a doubly tainted value to a singly tainted
  one.
%
  By contrast with the situation for abelian group indexed types, the
  lack of inverses in the monoid indexed setting means that the type
  $\forall m_1,m_2\mathord:\mathsf{M}.\ \tyPrim{M}{m_1m_2} \to
  \tyPrim{M}{m_2} \to \tyPrim{M}{m_1}$ is also uninhabited since there
  are no naturals $a_1,a_2$ such that $(m_1m_2)^{a_1}m_2^{a_2} \equiv
  m_1$.
\end{example}

>>>>

For the relational environments, we use %make use of 
the compositional
construction in \autoref{sec:constr-rel-env}, setting the chosen model
$\mathcal{M}$ to be the centre $Z(M)$ of $M$. % (recall that the
%centre of a monoid is the set of all elements that commute with every
%other element). 
Setting $R_{\tyPrimNm{M}}(m) = \{ (n,mn) \sepbar n
\in M \}$
%By setting the model $\mathcal{M}$ to be only the elements of $M$ that
%commute with everything else, we ensure that the following lemma
%holds:
gives:
\begin{lemma}\label{lem:monoid-ops-related}
  For all $\Delta$ and $\rho \in \relEnv{E}_\Mon(\Delta)$,
  $(\eta_\Mon,\eta_\Mon) \in \rsem{\Gamma_\Mon}{\relEnv{E}}\rho$.
\end{lemma}

@@

\begin{theorem}
  \label{thm:monoid-indefinability}
  Let $m$ and $n$ be natural numbers. For all $n$-by-$m$-matricies of
  natural numbers $A$ and $m$-by-$1$ matricies of natural numbers $b$,
  there exists an $M$ in the quantifier-free fragment, with
  \begin{equation}
    i_1,\dots,i_m\mathord:\mathsf{M}; \Gamma_\Mon \vdash M : \tyPrim{M}{e_1} \to \dots \to \tyPrim{M}{e_n} \to \tyPrim{M}{e}
  \end{equation}
  where $e_j = i_1^{A_{j1}}\dots i_m^{A_{jm}}$ and $e = i_1^{b_1}\dots
  i_m^{b_m}$ iff there is a solution $\vec{x} = (x_1,...,x_n)$ in the
  natural numbers to $A x^\top = b$.
\end{theorem}

The proof of this theorem uses the non-compositional relational
environments from \autoref{sec:constr-rel-env}, with
$R^\bullet_{\tyPrimNm{M}} = \emptyset$.

\begin{example}
  By \thmref{thm:monoid-indefinability}, it is impossible to write a
  program with the type $\tyPrim{M}{m_1m_2} \to \tyPrim{M}{m_1}$, and is thus % shows that it is
  impossible to convert a doubly tainted value to a singly tainted
  one.
%
  By contrast with the situation for abelian group indexed types, the
  lack of inverses in the monoid indexed setting means that the type
  $\tyPrim{M}{m_1m_2} \to \tyPrim{M}{m_2} \to \tyPrim{M}{m_1}$ is also
  uninhabited since there are no naturals $a_1,a_2$ such that
  $(m_1m_2)^{a_1}m_2^{a_2} \equiv m_1$.
\end{example}
